### 2012-11-01 counterexample to bratman on shared intention/a counterexample to bratman on shared intention.tex
working on it

`3. 1 and 2 are common knowledge between us' \citep[][p.\ View 4]{Bratman:1993je}
\end{quote}
%
Care is needed in specifying the contents of the intentions concerning our J-ing in these clauses. 
As mentioned above, an appeal to shared intention is supposed to  characterise systematically a difference between two agents who intentionally J together (e.g.\ walk  together) and two agents who also  J together but do not do so intentionally (e.g.\ they happen to be walking side by side). 
So we must avoid tacitly appealing to this distinction by  restricting possible values of J to those which involve our intentionally doing something together. 
Rather the above  conditions, (1)--(3), must be sufficient for shared intention even for some values of J which are `neutral with respect to shared intentionality'.%
\footnote{
 \citet[p.\ 147]{Bratman:1999fr}.
 This refines Bratman's earlier view that some admissable values of J are cooperatively neutral 
 	where an  act-type is \emph{cooperatively neutral} just if `joint performance of an act of that type may be cooperative, but it need not be' \citep[p.\ 330]{Bratman:1992mi}. 
}

A consequence is that, in the right situations, one of us can rationally intend that we J, and can intend this unilaterally, that is without depending on anyone else intending that we J. 
Suppose you know that I am going to Chicago via a certain route at a particular time, 
and that I will do this regardless of what you do.
Suppose also that you can rationally intend that you go to Chicago in the same manner.
Then you can rationally intend that we go to Chicago together, 
	and you can intend this irrespective of whether I have any corresponding intention---providing, of course, that in so intending you are conceiving of our going to Chicago together in a way that is neutral with respect to shared intentionality.
What follows depends on the premise that this is indeed possible.%
\footnote{ 
\citet{Bratman:1999fr}  defends this claim at length. 
Note also that this claim must be true if Bratman's account of shared intention is to provide an informative way of distinguishing between the contrast cases mentioned at the start.
}

@@

Before going further we must distinguish two versions of the claim that (1)--(3) are collectively sufficient for shared intention.
The \emph{weak claim} is that there is some J such that these conditions are sufficient for you and I to intend that we J.
The \emph{strong claim} is that for any J, these conditions are sufficient for you and I to intend that we J.
One of Bratman's aims is to show that it is possible to give an account of shared intention using concepts that `are available within the theory of individual planning agency' \citep[p.\ 163]{Bratman:2009lv}.  
Achieving this aim would require the strong claim, and it is to the strong claim that our counterexample is directed.

\section{Unshared Intentions}
Before we can introduce the counterexample we need to introduce what we shall call `unshared intentions'.
We have an \emph{unshared intention} that we <J$_1$, J$_2$> where J$_1$$\neq$J$_2$ just if:
%
\begin{quote}
\label{df:unshared_intention}
1. (a) I intend that we J$_1$ and (b) you intend that we J$_2$

@@

This is not quite enough to meet the second condition, (2), because there is also a requirement about meshing subplans. 
To make sense of this requirement we need to generalise Bratman's definition of meshing:
\begin{quote}
`our individual subplans concerning our J-ing \emph{mesh} just in case there is some way we could J that would not violate either of our subplans but would, rather, involve the successful execution of those subplans' \citep[p.\ 106]{Bratman:1993je}.
\end{quote}
A natural generalisation is this:
\begin{quote}
our individual subplans concerning our <J$_1$, J$_2$>-ing \emph{mesh} just in case there is some way we could <J$_1$, J$_2$> that would not violate either of our subplans but would, rather, involve the successful execution of those subplans. 
\end{quote}
%
To illustrate, 
there would be a failure to mesh if, in intending that they make the cross hit the red square,  Ayesha's plans had included pushing Benji out of the way and seizing his controls. 
But in the case we have been describing there is no such failure to mesh.
Each agent's subplans involve manipulating his or her own controls,
and the successes each seeks in doing this depends on the other successfully carrying out their subplans.
So Ayesha can rationally intend that they, Ayesha and Benji, make the cross hit the red square in accordance with and because of their intentions and meshing subplans of them.
And Benji likewise for making the cross hit the blue circle.

>>>>

`3. 1 and 2 are common knowledge between us' \citep[][p.\ View 4]{Bratman:1993je}
\end{quote}
%
How could we show that our meeting these conditions is not in fact sufficient for us to have a shared intention? 
We seek a case where the conditions are met but we lack a shared intention.
But, in advance of deciding whether Bratman's account is right, how could we determine whether we have a shared intention?
As already mentioned, an appeal to shared intention is supposed to  characterise systematically a difference between two agents who intentionally J together (e.g.\ walk  together) and two agents who also  J together but do not do so intentionally (e.g.\ they happen to be walking side by side). 
Let's stipulate that two cases \emph{differ with respect to shared agency} when they differ in the way exemplified by the contrast between our intentionally walking together and our merely walking in parallel.
Suppose, then, that we construct three cases, A, B and C.
The three cases are as similar as possible, except that the above conditions are met in cases B and C but not in case A.
Now if Bratman's conditions are sufficient for shared intention, A should differ from B with respect to shared agency, and B should not differ from C with respect to shared agency.
But suppose that in fact A does not differ from B with respect to shared agency, but B does so differ from C.%
\footnote{
Strictly speaking only A and B are needed for the counterexample, of course.
But including C makes it clearer that A and B do not differ with respect to shared agency.
}
Then we can conclude that Bratman's conditions are not sufficient for shared intention.
This is how our counterexample will work.

Care is needed in specifying the contents of the intentions concerning our J-ing in the above clauses, (1)--(3). 
As we have stressed, 
an appeal to shared intention is supposed to characterise systematically differences with respect to shared agency---to characterise, that is, a difference between two agents who intentionally J together and two agents who also  J together but do not do so intentionally.
We must therefore avoid tacitly appealing to this distinction by  restricting possible values of J to those which involve our intentionally doing something together. 
Rather the above  conditions, (1)--(3), must be sufficient for shared intention even for some values of J which are `neutral with respect to shared intentionality'.%
\footnote{
 \citet[p.\ 147]{Bratman:1999fr}.
 This refines Bratman's earlier view that some admissable values of J are cooperatively neutral 
 	where an  act-type is \emph{cooperatively neutral} just if `joint performance of an act of that type may be cooperative, but it need not be' \citep[p.\ 330]{Bratman:1992mi}. 
}

A consequence is that, in the right situations, one of us can rationally intend that we J, and can intend this unilaterally, that is without depending on anyone else intending that we J. 
Suppose you know that I am going to Chicago via a certain route at a particular time, 
and that I will do this regardless of what you do.
Suppose also that you can rationally intend that you go to Chicago in the same manner.
Then you can rationally intend that we go to Chicago together, 
	and you can intend this irrespective of whether I have any corresponding intention---providing, of course, that in so intending you are conceiving of our going to Chicago together in a way that is neutral with respect to shared intentionality.
What follows depends on the premise that this is indeed possible.%
\footnote{ 
\citet{Bratman:1999fr}  defends this claim at length. 
Note also that this claim must be true if Bratman's account of shared intention is to provide an informative and systematic distinction between the contrast cases mentioned at the start.
}

@@

Before going further we must distinguish two versions of the claim that (1)--(3) are collectively sufficient for shared intention.
The \emph{weak claim} is that there is some J such that these conditions are sufficient for you and I to intend that we J.
The \emph{strong claim} is that for any J, these conditions are sufficient for you and I to intend that we J.
One of Bratman's aims is to show that an account of shared intention can be \emph{conceptually conservative}:
he aims, that is, to show that it is possible to give an account of shared intention using concepts that `are available within the theory of individual planning agency' \citep[p.\ 163]{Bratman:2009lv}.  
Achieving this aim would require the strong claim, and it is to the strong claim that our counterexample is directed.

\section{Unshared Intentions}
Before we can introduce the counterexample we need to define what we shall call `unshared intentions'.
We have an \emph{unshared intention} that we <J$_1$, J$_2$> where J$_1$$\neq$J$_2$ just if:
%
\begin{quote}
\label{df:unshared_intention}
1. (a) I intend that we J$_1$ and (b) you intend that we J$_2$

@@

This is not quite enough to meet the second condition, (2), because there is also a requirement about meshing subplans. 
To make sense of this requirement we need to generalise Bratman's definition of meshing:
\begin{quote}
`our individual subplans concerning our J-ing \emph{mesh} just in case there is some way we could J that would not violate either of our subplans but would, rather, involve the successful execution of those subplans' \citep[p.\ 106]{Bratman:1993je}.
\end{quote}
A natural generalisation is this:
\begin{quote}
our individual subplans concerning our <J$_1$, J$_2$>-ing \emph{mesh} just in case there is some way I could J$_1$ and you could J$_2$ that would not violate either of our subplans but would, rather, involve the successful execution of those subplans. 
\end{quote}
%
To illustrate, 
there would be a failure to mesh if, in intending that they make the cross hit the red square,  Ayesha's plans had included pushing Benji out of the way and seizing his controls. 
But in the case we have been describing there is no such failure to mesh.
Each agent's subplans involve manipulating his or her own controls,
and the successes each seeks in doing this depends on the other successfully carrying out their subplans.
So Ayesha can rationally intend that they, Ayesha and Benji, make the cross hit the red square in accordance with and because of their intentions and meshing subplans of them.
And Benji likewise for making the cross hit the blue circle.
